---
title: "FM_Assignment7_Duo_Zhou"
author: "Duo Zhou"
date: "8/10/2020"
output: pdf_document
---

```{r, include=FALSE}
suppressWarnings(library(RQuantLib))
suppressWarnings(library(fOptions))
```

### Loading Data
```{r}
datapath <- "C:/Users/zd000/Desktop/MSCA/Financial Analytics/Assignments/week7/"
optionsData = read.csv(paste(datapath,'OptionsPrices.csv',sep='/'),header=TRUE,row.names=1,sep=",")
dim(optionsData) 
head(optionsData,10)
```

### Using the options data from file OptionsPrices.csv used in the workshop calculate the P&L breakdown into delta, gamma and vega components of 1425 December 2007 put for each day between 2/27/2007 and 3/5/2007.
### Compare actual P&L for 1425 December 2007 put with the estimated daily P&L for each day between 2/27/2007 and 3/5/2007.
### Put the results in a table with columns: actual P&L, estimated P&L, delta P&L, gamma P&L, vega P&L.

First, Calculate implied volatilites of put with strike 1425 between 2/26/2007 and 3/5/2007

```{r}
# Use europeanoptionimplied vol() over each row of optionsData, 
## use as.vector to eliminate dates and multiply by 100 to show in percent
put1425.impliedVolatilities <- as.vector(apply(optionsData[1:6,], 1, function (z)
EuropeanOptionImpliedVolatility("put", value=z[26], underlying=z[3], strike=1425,
dividendYield = z[2], riskFreeRate = z[2], maturity = z[1]/365, volatility = 0.3)))*100
# return head of implied 1425p vols
cbind(row.names(optionsData[1:6,]), put1425.impliedVolatilities)
```


Find option premimum and Greeks of 1425 put between 2/26/2007 and 3/5/2007
```{r}
currentPutValue<-matrix(NA, nrow = 6, ncol = 4)
for(i in 1:6){
currentPutValue.temp <- EuropeanOption(type="put",underlying = optionsData[i,3],
strike = 1425, maturity = optionsData[i,1]/365, riskFreeRate = optionsData[i,2],
volatility =put1425.impliedVolatilities[i]/100 , dividendYield = optionsData[i,2])
currentPutValue[i,1]<-as.numeric(currentPutValue.temp[1])
currentPutValue[i,2]<-as.numeric(currentPutValue.temp[2])
currentPutValue[i,3]<-as.numeric(currentPutValue.temp[3])
currentPutValue[i,4]<-as.numeric(currentPutValue.temp[4])
}
colnames(currentPutValue)<-c('value','delta','gamma','vega')
rownames(currentPutValue)<-row.names(optionsData[1:6,])
# Return the value and greeks
currentPutValue
```

Calculate the daily breakdown of 1425-put P&L between 2/27/07 and 3/05/07 into delta gama and vega buckets
First find estimated delta, gamma and vega P&L using forula on slide 15 for this week and then sum them up to get total estimated P&L. The actual P&L is calculated by taking the difference between consecutive daily P1425 premimum.
```{r}
PNL.daily<-matrix(NA, nrow = 5, ncol = 5)

for(i in 1:5){
# delta pnl= delta* change in price
PNL.daily[i,3]<-currentPutValue[i,2]*(optionsData[i+1,3]-optionsData[i,3])
# gamma pnl= (1/2) * gamma * (change in price)^2
PNL.daily[i,4]<-.5*currentPutValue[i,3]*(optionsData[i+1,3]-optionsData[i,3])^2
# vega pnl= (vega/100) * change in implied volatility
PNL.daily[i,5]<-currentPutValue[i,4]/100*
  (put1425.impliedVolatilities[i+1]-put1425.impliedVolatilities[i])

# Calculate the sum of the greek pnls
PNL.daily[i,2]<-PNL.daily[i,3]+PNL.daily[i,4]+PNL.daily[i,5]
# Calculate the actual put pnl
PNL.daily[i,1]<-optionsData[i+1,26]-optionsData[i,26]
}
colnames(PNL.daily)<-c('actual P&L', 'estimated P&L', 'delta P&L', 'gamma P&L',
                             'vega P&L')
rownames(PNL.daily)<-row.names(optionsData[2:6,])
PNL.daily
```
