---
title: "DM_Assignment4_Data_PrePorcessing"
author: "Duo Zhou"
date: "8/3/2020"
output: html_document
---
### Pre-Process Raw Data for this Assignment

```{r}
dataPath <- "C:/Users/zd000/Desktop/MSCA/Data Mining/Assignments/week4/"
DB <-read.csv(paste(dataPath, 'diabetes_data.csv', sep = '/'), header = TRUE)
# remove Categorical Variables, keeping only the numerical data 
head(DB)
```

```{r, include=FALSE}
replace(DB, DB =="?", NA) # replace '?' to NA to understand missing value better
```

### Diag_1:
The dataset contained upto three diagnoses for a given patient (primary (diag_1), secondary(diag_2) and additional(diag_3)). However, each of these had 700–900 unique ICD codes
You can collapsed these diagnosis codes into 9 disease categories . These 9 categories include Circulatory, Respiratory, Digestive, Diabetes, Injury, Musculoskeletal, Genitourinary, Neoplasms, and Others.

You can convert all 3 into 9 categories each. however primary diagnosis is enough for this study.It is upto you!!
ICD9 code conversion reference (Read about it online to understand more)

ICD-9 CM Category ICD-9 CM Codes Frequency % of Total Circulatory System 390-459.9 562 30.6% Respiratory System 460-519.9 258 13.8% Digestive System 520-579.9 223 12.2% Neoplasms 140-239.9 123 6.7% Genitourinary System 580-629.9 114 6.2% Signs, Symptoms, and 780-799.9 103 5.6% Ill-Defined Conditions Mental Disorders 290-319 85 4.6% Musculoskeletal System 710-739.9 83 4.5% and Connective Tissue Injury and Poisoning 800-999.9 83 4.5% Endocrine, Nutritional, 240-279.9 69 3.8% Metabolic ...

```{r}
convert_diag_codes<- function(code){
    if(is.null(code)){
        return('Other')}
    else if(grepl('V' , code) | grepl('E' , code)) {
        return ('Other')}
    else{
        code = as.numeric(code)
        if ((code >= 390) & (code < 460) | (floor(code)==785)){
            return('Circulatory')}
        else if ((code >= 460) & (code < 520) | (floor(code)==786)){
            return('Respiratory')}
        else if ((code >= 520) & (code < 580) | (floor(code)==787)){
            return('Digestive')}
        else if (code == 250){
            return('Diabetes')}
        else if ((code >= 800) & (code < 1000)){
            return('Injury')}
        else if ((code >= 710) & (code < 740)){
            return('Musculoskeletal')}
        else if ((code >= 580) & (code < 630) | (floor(code)==788)){
            return('Genitourinary')}
        else if ((code >= 140) & (code < 240)){
            return('Neoplasms')}
        else{
            return('Other')}
    }
  
}

```

```{r}

DB['diag_1']<-apply(DB['diag_1'],1,convert_diag_codes)

```

### Convert age to Integer

```{r}
library(plyr)
DB$age<-mapvalues(DB$age, 
    c("[0-10)", "[10-20)", "[20-30)", "[30-40)", "[40-50)",
"[50-60)", "[60-70)", "[70-80)", "[80-90)", "[90-100)"), 
c(5,15,25,35,45,55,65,75,85,95))
DB['age']<-apply(DB['age'],1,as.integer)
```

### Convert 2 value categories into binary variable

```{r}
DB$change<-mapvalues(DB$change, c('Ch','No'),c(1,0))
DB$gender<-mapvalues(DB$gender, c('Male','Female'),c(1,0))
DB$diabetesMed<-mapvalues(DB$diabetesMed, c('Yes','No'),c(1,0))
```

### Merge categories to one/two level less - Use domain knowledge
```{r}
DB$A1Cresult<-mapvalues(DB$A1Cresult, c('>7','>8','Norm','None'),
                        c('Abnormal','Abnormal','Normal','Not tested'))
DB$max_glu_serum<-mapvalues(DB$max_glu_serum, c('>200','>300','Norm','None'),
                        c('Abnormal','Abnormal','Normal','Not tested'))
```

### Merge some categories together - By studying other variables

Admission source, admission type and discharge disposition can help us to understand what theses admission type id corresponds to. For example, admission types 1, 2 and 7 correspond to Emergency, Urgent Care and Trauma, and thus were combined into a single category as these are all non-elective situations.

```{r}
DB$admission_type_id<-mapvalues(DB$admission_type_id, c(1,2,3,4,5,6,7,8),
                        c(1,1,3,4,5,5,1,5))
```

### Some columns looks numerical but they are not... Convert them into string
example: admission_type_id, discharge_disposition_id

```{r}
DB['admission_type_id']<-apply(DB['admission_type_id'],1,as.character)
DB['discharge_disposition_id']<-apply(DB['discharge_disposition_id'],1,as.character)
```

### Remove duplicate patients; 
-Either simply remove them
-Find a way to deal with them - domain knowledge

```{r}
DB<-DB[!duplicated(DB$patient_nbr), ]

```

### Drop columns that are not needed for the analysis


```{r}
DB<-subset(DB, select = -c(patient_nbr,diag_2,diag_3,encounter_id,
                       admission_source_id))
```

### Check missing value

```{r}
miss.value.percent<-function(df){
   MV.per<- data.frame()
   for (cn in colnames(df)){
      MV.per[cn,'MV Percentage']<-sum(is.null(DB[cn]))/dim(DB)[1]*100
}
return(MV.per)
}

miss.value.percent(DB)
```

No missing values!

### Change readmitted (response variable) into binary variable
```{r}
DB$readmitted<-mapvalues(DB$readmitted, c('>30','<30','NO'),c(1,1,0))
head(DB)
```

Remove columns with ony 1 factor
```{r}
onefactor<-c('acetohexamide','glimepiride.pioglitazone')
DB<-DB[,!(names(DB) %in% onefactor)]
dim(DB)
```


```{r}
write.csv(DB,paste(dataPath,'diabetes_data_preprocess.csv'), row.names=FALSE)  

```

SAVE Train and Test Data in csv file for model comparison between different methods.

```{r}
train_size <- floor(0.7 * nrow(DB))
set.seed(12345)
# seperate the data into training and test set
train_ind <- sample(nrow(DB), size = train_size)
DB.train <- DB[train_ind, ]
DB.test <- DB[-train_ind, ]
dim(DB)
dim(DB.train)
dim(DB.test)
write.csv(DB.train,paste(dataPath,'DB_train.csv'), row.names=FALSE)
write.csv(DB.test,paste(dataPath,'DB_test.csv'), row.names=FALSE)
```


